Efficient Selection of Time Series Shapelets

### 2024-04-19 Docs/paper/main.tex
Clustering based shapelets, example and some questions

\author{Adam Charane}
\affiliation{%
	\institution{Free University of Bozen-Bolzano}
	\city{Bolzano}
	\state{Italy}
}
\email{acharane@unibz.it}

\author{Matteo Ceccarello}
\affiliation{%
	\institution{University of Padova}
	\city{Padova}
	\state{Italy}
}
\email{matteo.ceccarello@dei.unipd.it}

\author{Johann Gamper}
\affiliation{%
	\institution{Free University of Bozen-Bolzano}
	\city{Bolzano}
	\state{Italy}
}
\email{johann.gamper@unibz.it}

@@

\subsection{Clustering-based shapelets}
\begin{figure*}
  \includegraphics[width=\textwidth]{imgs/clustering.png}
  \caption{Sliding windows of length 40 from \emph{CBF} dataset, with centers
  (red dots) of 100 clusters using K-Means algorithm.\label{fig:clustering}}
\end{figure*}
We start by extracting the set of sliding windows $W$ from all the time series
$W = \bigcup_T W_T^l$, then we cluster all the windows in $W$.
For each cluster $C_i$, we assign a label $c$ using a majority vote, i.e., the
label of cluster $C_i$ is the label of the set with largest cardinal, consisting
of windows assigned to cluster $C_i$ that have the same class label.
Figure~\ref{fig:clustering} shows an example from the \emph{CBF} dataset plotted
on the plan by keeping only the first two PCA components. The gray dots in the
left sub-figure are the windows of length 40, and the red dots are the centers
of the clusters obtained y running K-Means algorithm with 100 clusters. The
right sub-figure shows the labels of the subsequences.

>>>>

%\author{Adam Charane}
%\affiliation{%
%\institution{Free University of Bozen-Bolzano}
%\city{Bolzano}
%\state{Italy}
%}
%\email{acharane@unibz.it}

%\author{Matteo Ceccarello}
%\affiliation{%
%\institution{University of Padova}
%\city{Padova}
%\state{Italy}
%}
%\email{matteo.ceccarello@dei.unipd.it}

%\author{Johann Gamper}
%\affiliation{%
%\institution{Free University of Bozen-Bolzano}
%\city{Bolzano}
%\state{Italy}
%}
%\email{johann.gamper@unibz.it}

@@

\subsection{Clustering-based shapelets}
\begin{figure*}
	\includegraphics[width=\textwidth]{imgs/clustering.png}
	\caption{Sliding windows of length 40 from \emph{CBF} dataset, with centers
		(red dots) of 100 clusters using K-Means algorithm.\label{fig:clustering}}
\end{figure*}
We start by extracting the set of sliding windows $W$ from all the time series
$W = \bigcup_T W_T^l$, then we cluster all the windows in $W$.
For each cluster $C_i$, we count the fraction of windows from each class
(popularity), and only keep clusters where the popularity of one class is beyond
a threshold. Figure~\ref{fig:clustering} shows an example from the \emph{CBF}
dataset plotted on the plan by keeping only the first two PCA components. The
gray dots in the left sub-figure are the windows of length 40, and the red dots
are the centers of the clusters obtained y running K-Means algorithm with 100
clusters. The right sub-figure shows the labels of the subsequences.
Figure~\ref{fig:clustering_threshold} shows the same windows, but this time
after removing all the clusters where the popularity of the dominant class is
below some certain threshold.
The idea is that we want to get rid of all subsequences from different classes
that are similar to each other, since a shapelet should be a pattern repeated
between time series from the same class.
\begin{figure*}
	\includegraphics[width=\textwidth]{imgs/clustering_thresholds.png}
	\caption{Windows left after removing all the clusters where the popularity of
		the dominant class is below a threshold\label{fig:clustering_threshold}.}
\end{figure*}

\ac{
	There are many questions/decisions that have to be made:
	\begin{itemize}
		\item How many clusters we should use?
		\item Which clusters to keep (so far I am using the popularity)?
		\item Should we use all the clusters kept?
	\end{itemize}
	Some information that we have and that have not been used yet:
	\begin{itemize}
		\item In each cluster, besides the popularity of each class, we know also
		      how many time series are covered. For example, in the \emph{CBF}
		      dataset, we have 30 training examples, 10 for each class. If we have
		      two clusters with popularity, let's say 90\% for class 1, the windows
		      from cluster 1 are all extracted from 2 time series, and in the second
		      clusters, the windows are from all 10 time series, clearly we have to
		      favor the second cluster, since it has a pattern that is repeated
		      between all the of the time series from the same class. How can we
		      use this information?
		\item How to use the distance between the centers of the clusters?
	\end{itemize}
}
